\documentclass[12pt]{article}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{pifont}
\usepackage{natbib}
\usepackage{color}
\def\bibfont{\footnotesize}
\setlength{\bibsep}{0pt plus 0.3ex}

\usepackage{url}

\begin{document}
\def\R{\mathbb{R}}
\def\C{\mathbb{C}}
\def\Z{\mathbb{Z}}
\def\N{\mathbb{N}}
\def\Q{\mathbb{Q}}
\def\D{\mathbb{D}}
\def\Sp{{\mathbb{S}}}
\def\T{\mathbb{T}}
\def\H{\mathbb{H}}
\def\hb{\hfil \break}
\def\ni{\noindent}
\def\i{\indent}
\def\a{\alpha}
\def\b{\beta}
\def\e{\epsilon}
\def\d{\delta}
\def\D{\Delta}
\def\G{\Gamma}
\def\g{\gamma}
\def\l{\lambda}
\def\m{\mu}
\def\s{\sigma}
\def\Si{\Sigma}
\def\th{\theta}
\def\z{\zeta}
\def\p{\partial}
\def\o{\omega}
\def\O{\Omega}
\def\t{\tau}
\def\L{\it \char'44}
\def\F{\mathcal{F}}
\def\B{\mathcal{B}}
\def\C{\mathcal{C}}
\def\half{\frac{1}{2}}
\ni m3f33prob2 \\
\begin{center}
{\bf M3F22 PROBLEMS 2.  20.10.2017} 
\end{center}
\i In all Applied Mathematics, it is important to have a reasonable grasp of the subject area to which mathematics is to be applied.  Here in Ch. II, this is economic and financial -- with inescapable political overtones.  Accordingly, the questions below do not have definitive answers.  They are put to encourage you to think about underlying issues, and to do a little digging in the Internet.  The corresponding Solutions 2 consist largely of edited extracts from Wikipedia. \\

\ni Q1: {\it Keynes and Hayek/Friedman}. \\  
\i The two most influential economists of the last century were Keynes and (according to taste) one of Hayek and Friedman.  Give brief accounts of their work, and of their influence on policy. \\

\ni Q2: {\it ``Physics and Barings"}.  \\
\i Justify the statement in the text, that it was obviously impossible to systematically exploit price movements
between these two exchanges in either direction. \\

\i Other questions of this kind (M3F22 2015-16, Problems 1, 2) include subjects such as: LTCC; the Euro; interest rates; quantitative easing (QE); the People's Republic of China (economic policy and implications here); large-scale failures; sovereign default; large-scale fraud; systemic instability.  See last year's website for anything you wish to pursue.  Even better, think up things you would like to pursue for yourself, and do so. \\

\hfil NHB \break

\end{document}

\ni Q2.  One of the great problems in the financial system is {\it instability} (cf. the meeting on 30.9.2015, CEPR Financial Regulation Initiative), with great knock-on effects on the real economy, job prospects (including yours), etc.  Do you have a view as to what can/should/might be done? \\
